Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions

### 2010-11-09 rjournal/RJtemplate.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@22 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\abstract{
   A general theory for extending nonparametric goodness-of-fit tests to discrete null distributions has existed
   for several decades. Despite this, modern statistical software have generally failed to provide this methodology
   to users. The packages ks.test and cvm.test serve to fill this need in the R language for
   the two most popular nonparametric tests. The structure of these two packages are explained and examples
   to their specific usage are presented. Particular attention is given to the various numerical issues that arise
   in their implementation.
}

@@

%Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
While almost any hypothesis test can be viewed as a variant on a goodness-of-fit test, the term is typically
applied to those tests which are nonparametric in nature. That is, they do not conduct the hypothesis testing 
through the explicit calculation of the parameters generating the data. Instead, they generally attempt to 
calculate differences between the empirical distribution of the observed data and the overall distribution 
of the null model. In many cases these tests are preferential since they tend to have an increased power for 
interesting deviations from the null model, in exchange for failing to detect less interesting deviations due to 
factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r and von Mises.

@@

Hmm... I thought the package had a trivial example where the standard
\code{ks.test()} was a problem?
\begin{Schunk}
\begin{Sinput}
> library(ks.test)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7  4  8  5  8 10
[19]  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, p-value = 1
alternative hypothesis: two-sided 
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, p-value = 0.7731
alternative hypothesis: the CDF of x lies above the null hypothesis 
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g", exact = FALSE)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

Discuss this, nice example.  Relate to new material that is about to go into
the earlier section.
\begin{Schunk}
\begin{Sinput}
> stats::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  c(0, 1) 
D = 0.5, p-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\begin{Sinput}
> ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

JAY: simulation showing original ks.test() was conservative, Conover exact.

>>>>

\abstract{
A general theory for extending nonparametric goodness-of-fit tests 
to discrete null distributions has existed
for several decades. Despite this, modern statistical software 
has generally failed to provide this methodology
to users. We offer a revision of R's \code{ks.test()} function
and a new \code{cvm.test()} function that serve to fill this need
in the R language for two of the most popular nonparametric tests. 
This paper describes the the structure of these contributions and
provides examples of their usage. Particular attention is given
to various numerical issues that arise in their implementation.
}

@@

%Why we use non-parametric tests (power to interesting shifts; robustness to underlying assumptions).
While almost any hypothesis test can be viewed as a variant on a
goodness-of-fit test, the term ``goodness-of-fit test'' is typically
reserved for those tests which are nonparametric in nature. 
That is, they do not conduct the hypothesis testing 
through the explicit calculation of the parameters generating the data. Instead, they generally attempt to 
calculate differences between the empirical distribution of the observed data and the overall distribution 
of the null model. In many cases these tests are preferential since they tend to have an increased power for 
interesting deviations from the null model, in exchange for failing to detect less interesting deviations due to 
factors such as measurement error. By far the most popular of these nonparametric tests are due to Kolmogorov and 
Smirnov, followed closely by several variants on an estimation procedure proposed by Cram\'{e}r and von Mises.

@@

Consider a toy example, with observed data of length 2 (specifically, the
values 0 and 1) and a hypothized null distribution that places equal
probability on the values 0 and 1.  With the current \code{ks.test()} function
in R (which, admittedly, doesn't claim to handle discrete distributions),
the reported p-value, 0.5, is clearly incorrect:
\begin{Schunk}
\begin{Sinput}
> stats::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  c(0, 1) 
D = 0.5, p-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Instead, the value of $D$ given in equation (1)
should be 0 and the associated p-value should be 1.  Our revision of \code{ks.test()}
fixes this problem when the user provides a discrete distribution:
\begin{Schunk} 
\begin{Sinput}
> ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

Next, we simulate a sample of 25 from the discrete uniform distribution on
the integers $\{1, 2, \ldots, 10\}$ and show several variants of the
new \code{ks.test()} implementation.  The first is the default two-sided test,
where the reported p-value is a conservative upper bound for the actual
p-value.  In this case, the approximation may not be that tight, but this is
irrelevant for such large p-values (for more interesting p-values, the
upper bound is very close to the true p-value).
\begin{Schunk}
\begin{Sinput}
> library(ks.test)
> set.seed(1)
> x <- sample(1:10, 25, replace = TRUE)
> x
\end{Sinput}
\begin{Soutput}
 [1]  3  4  6 10  3  9 10  7  7  1  3  2  7
[14]  4  8  5  8 10  4  8 10  3  7  2  3
\end{Soutput}
\begin{Sinput}
> ks.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D = 0.08, p-value = 1
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Next, we conduct the default one-sided test, where Conover's method
provides the exact p-value (up to the numerical precision of the
implementation):
\begin{Schunk}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g")
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

data:  x 
D^+ = 0.04, p-value = 0.7731
alternative hypothesis:
the CDF of x lies above the null hypothesis 
\end{Soutput}
\end{Schunk}
In contrast, the option \code{exact=FALSE} results in the
p-value obtained by applying the classical Kolmogorov-Smirnov
test, resulting in a conservative p-value:
\begin{Schunk}
\begin{Sinput}
> ks.test(x, ecdf(1:10), alternative = "g", 
+         exact = FALSE)
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

JAY: reconsider simulation.  Not really needed, if we have citations for
the conservatism points, etc...?  The code for this is in the JSS folder.
if we want to revisit it.

Finally, we employ two of the Cram\'{e}r-von Mises tests.  Taylor, are these
your ``cynical example'' that you commented on, or did you have something
in addition to these?  What should be discussed here?
